---
title: "Aggregating anemone data analysis"
author: "Caitlin Bergman and Beth Blanchette"
date: "11/11/2021"
output:
  pdf_document: default
  html_document: default
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

## Loading libraries and data
```{r, message = F, warning = F}
#loading libraries
library(rstatix)
library(fitdistrplus) 
library(patchwork)
library(ordinal)
library(performance)
library(scales)
library(gamlss)
library(brms) 
library(tidyverse)

#Loading data
pam<-read_delim("data/pam_data_clean.csv", delim = ",")
base <- read_delim("data/base_diameter_data_clean.csv", delim = ",")
food <- read_delim("data/feeding_time_data_clean.csv", delim = ",")
open_closed <- read_delim("data/open_closed_data_clean.csv", delim = ",")
hemocytometer <- read_delim("data/hemocytometer_data_clean.csv", delim = ",")
```

# Cleaning data
```{r, message = F, warning = F}
#Creating a clean dataframe for each response variable, and a summarized dataframe with average #value for each treatment at each measurement time

#Function to calculate standard error for data summaries
standard_error <- function(x) sd(x) / sqrt(length(x))

#Cleaning photosynthetic efficiency data: Filtering to remove measurements that were not used 
#during analysis, formatting columns, and selecting columns needed for model.
pam_clean <- pam %>%
  filter(Measurement != "NA")%>% 
  select(Date,Event,Measurement,Treatment,Bin,Site,Anemone_ID,Fv_Fm_av)%>%
  mutate(Date = as.POSIXct(as.character(Date),
                           format="%m/%d/%Y %H:%M:%S"),
         Anemone_ID = as.factor(Anemone_ID), 
         Bin = as.factor(Bin), 
         Measurement = as.factor(Measurement), 
         Site = as.factor(Site), 
         Event = factor(Event, levels = c("Acclimation","Pre-heat", "Post-heat", "Recovery")),
         Treatment = fct_relevel(as.factor(Treatment), "Control", "25C", "30C"))

#Mean and standard error of photosynthetic efficiency for each treatment and measurement time
pam_summary <- pam_clean%>%
  group_by(Date, Measurement, Treatment) %>%
  summarize(mean_FvFm = mean(Fv_Fm_av), 
            se_FvFm = standard_error(Fv_Fm_av))

#Cleaning base measurement data: Filtering to remove measurements that were not used during 
#analysis, formatting columns, and selecting columns needed for model.
base_clean<-base%>%
  filter(Treatment != "NA", 
         Average_Diameter != "NA")%>% 
  mutate(Date = factor(Date, levels =c("31-Oct", "05-Nov", "09-Nov", "13-Nov")), 
         Event = fct_relevel(as.factor(Event), "Acclimation","Before heatwave", "After heatwave", "Recovery"),
         Treatment = fct_relevel(as.factor(Treatment), "Control", "25C", "30C"), 
         Anemone_ID = as.factor(Anemone_ID), 
         Bin = as.factor(Bin), 
         Site = as.factor(Site)) %>%
select(Date,Event,Treatment,Bin,Site,Anemone_ID,Average_Diameter)%>%
   arrange_all()

#Mean and standard error of base measurement data for each treatment and measurement time
base_summary <- base_clean%>%
  group_by(Date, Treatment)%>%
  summarize(mean_base = mean(Average_Diameter), 
            se_base = standard_error(Average_Diameter))

#Cleaning feeding time data: Filtering to remove measurements that were not used during 
#analysis, formatting columns, and selecting columns needed for model.
food_clean <- food%>%
  filter(Date != "10/28/2021")%>%
  mutate(Feeding_Time_Min = as.numeric(Feeding_Min),
         Event = fct_relevel(as.factor(Event),"Acclimation", "Before heatwave", "After heatwave", "Recovery"),
         Date = as.factor(Date),
         Site = as.factor(Site),
         Treatment = fct_relevel(as.factor(Treatment), "Control", "25C", "30C"), 
         Anemone_ID = as.factor(Anemone_ID), 
         Bin = as.factor(Bin))%>%
select(Date, Event,Treatment, Bin, Site, Anemone_ID, Feeding_Time_Min)

#Mean and standard error of feeding time data for each treatment and measurement time
food_summary <- food %>%
  group_by(Date, Treatment) %>%
  summarize(mean_time = mean(Feeding_Time_Min), 
            se_time = standard_error(Feeding_Time_Min))

#Cleaning heatwave response data: Formatting columns and selecting columns needed for model.
open_closed_clean <- open_closed %>%
  mutate(Date = as.factor(Date),
         Anemone_ID = as.factor(Anemone_ID),
         Time_Block = fct_relevel(as.factor(Time_Block), "0", "1", "2", "3", "4", "5", "6"), 
         Treatment = as.factor(Treatment),  
         Open_Closed = as.factor(Open_Closed), 
         Open_Closed = factor(Open_Closed, levels = c("Open", "Partially open", "Closed"), ordered = TRUE), 
         Treatment = fct_relevel(Treatment, "Control", "25C", "30C"))%>%
  select(Date, Event, Time_Block,Bin, Treatment, Open_Closed, Anemone_ID)

#Counts of heatwave response data for each treatment and measurement time
open_closed_summary <- open_closed_clean %>%
group_by(Date, Event, Treatment, Time_Block)%>%
  count(Open_Closed)
  
#Cleaning hemocytometer data: Converting units for mass to mg, calculating cell densities and 
#mitotic index, formatting columns, and selecting columns needed for model. 
hemo_clean <- hemocytometer %>%
  mutate(Tentacle_Mass_mg = (Tentacle_Mass_g*1000), 
         Dino_Density = ((Number_Dino_Average*0.5)/(Tentacle_Mass_mg*0.0001)), 
         Green_Density = ((Number_Green_Average*0.5)/(Tentacle_Mass_mg*0.0001)), 
         Dino_MI = (Dividing_Dino_Average/Number_Dino_Average)) %>% 
  mutate(Date = as.factor(Date), 
         Treatment = as.factor(Treatment), 
         Bin = as.factor(Bin), Site = as.factor(Site), 
         Anemone_ID = as.factor(Anemone_ID)) %>%
mutate(Date = as.POSIXct(as.character(Date),format="%m/%d/%Y"))%>%
  select(Date,Treatment, Bin, Site, Anemone_ID, Tentacle_Mass_mg, Number_Dino_Average, 
         Number_Green_Average, Dividing_Dino_Average, Dividing_Green_Average, Dino_Density, 
         Green_Density, Dino_MI) %>%
  group_by(Date, Treatment)

#Mean and standard error of cell density and mitotic index for zooxanthellae and zoochlorellae 
#at each treatment and measurement time
hemo_summary <- hemo_clean%>%
  group_by(Date, Treatment) %>%
  summarize(mean_Dino_Density = mean(Dino_Density), 
            se_Dino_Density = standard_error(Dino_Density), 
            mean_Green_Density = mean(Green_Density), 
            se_Green_Density = standard_error(Green_Density), 
            mean_Dino_MI = mean(Dino_MI), 
            se_Dino_MI = standard_error(Dino_MI))

#Summarizing mean and standard error of temperature data from heatwave
temp_summary <- open_closed %>%
  select(Date, Time_Block,Event, Treatment, Bucket_Temp) %>%
  group_by(Date, Event, Treatment, Time_Block) %>%
  summarize(mean_temp = mean(Bucket_Temp), 
            se_temp = standard_error(Bucket_Temp))%>%
  mutate(Date = as.factor(Date), 
         Treatment = as.factor(Treatment)) %>%
  mutate(Treatment = fct_relevel(Treatment, "Control", "25C", "30C"))
```

# PAM data analysis

## Plots
Plotting a timeseries including all photosynthetic efficiency measurement times
```{r, message = F, warning = F}
#Plotting measurements as a timeseries with standard error bars. 
ggplot(data = pam_summary, 
       aes(x=Date, 
           y = mean_FvFm, 
           group = Treatment, 
           colour = Treatment)) +
  theme_classic() +
  geom_vline(xintercept = as.POSIXct("2021-11-06 09:00:00"),
             linetype = "dotted", 
             size = 1) +
  geom_vline(xintercept = as.POSIXct("2021-11-08 16:00:00"), 
             linetype = "dotted", 
             size = 1)+
  geom_point(size=2.5) +
  geom_line(lwd = 1.5) +
  geom_errorbar(aes(ymin = mean_FvFm - se_FvFm, 
                    ymax = mean_FvFm + se_FvFm), 
                    width=30000)+
  labs(x="Date", 
       y = "Photosynthetic efficiency (Fv/Fm)") +
  scale_fill_manual(values =c('#89226AFF','#56B4E9FF','#E65154FF')) +
  scale_colour_manual(values = c('#89226AFF','#56B4E9FF','#E65154FF'))+
  theme(axis.text=element_text(size=15),
        axis.title=element_text(size=20), 
        legend.text=element_text(size=15), 
        legend.title=element_text(size=20))

ggsave(path = "plots",filename = "pam_overall_line.png", width = 10, height = 7)
```

Plotting a timeseries of all morning photosynthetic efficiency measurements (5 timepoints)
```{r, message = F, warning = F}
#Selecting all morning PAM measurements
pam_morning <- pam_summary%>%
  filter(Date == "2021-11-01 06:00:00"| 
         Date == "2021-11-06 06:00:00"| 
         Date == "2021-11-08 06:00:00"|
         Date == "2021-11-09 06:00:00"| 
         Date =="2021-11-13 06:00:00")

#Plotting measurements as a timeseries with standard error bars. 
ggplot(data = pam_morning,
       aes(x=Date, 
           y = mean_FvFm, 
           group = Treatment, 
           colour = Treatment)) +
  theme_classic() +
  geom_vline(xintercept = as.POSIXct("2021-11-06 09:00:00"),
             linetype = "dotted", 
             size = 1) +
  geom_vline(xintercept = as.POSIXct("2021-11-08 16:00:00"), 
             linetype = "dotted", 
             size = 1)+
  geom_point(size=3) +
  geom_line(lwd = 2) +
  scale_x_datetime(breaks=date_breaks("3 days"),labels=date_format("%b-%d"))+
  geom_errorbar(aes(ymin = mean_FvFm - se_FvFm, 
                    ymax = mean_FvFm + se_FvFm), 
                    width=30000)+
  labs(x="Date", 
       y = "Photosynthetic efficiency (Fv/Fm)") +
  scale_fill_manual(values =c('#89226AFF','#56B4E9FF','#E65154FF')) +
  scale_colour_manual(values = c('#89226AFF','#56B4E9FF','#E65154FF'))+
  theme(axis.text=element_text(size=15),
        axis.title=element_text(size=20), 
        legend.text=element_text(size=15), 
        legend.title=element_text(size=20))

ggsave(path = "plots",filename = "pam_timeseries.jpg", width = 15, height = 7)
```


Plot comparing morning and afternoon photosynthetic efficiency measurements on first and third days of heatwave:
```{r, message = F, warning = F}
#Selecting only PAM measurements taken on first and third heatwave days (Nov. 6 and Nov. 8)
pam_heatwave<-pam_clean%>%
  filter(Date == "2021-11-06 06:00:00"| 
        Date == "2021-11-06 16:00:00"| 
        Date == "2021-11-08 06:00:00"| 
        Date == "2021-11-08 16:00:00")%>%
  separate(Date, c('Day', 'Hour'), sep = " ", remove = T)%>%
 mutate(Day = fct_relevel(Day, "2021-11-06", "2021-11-08"))

#Changing names of heatwave days for x-axis of graph
  levels(pam_heatwave$Day) <- c("Heatwave day 1", "Heatwave day 3")

#Plotting data as two boxplots, separated by day. Each boxplot is separated by time (before/after heatwave) and treatment.
ggplot(pam_heatwave, 
       aes(fill=Treatment, 
           y=Fv_Fm_av, 
           x=Event)) + 
  geom_boxplot() +
 scale_fill_manual(values =c('#89226AFF','#56B4E9FF','#E65154FF')) +
   labs(x = "Measurement time", 
        y = "Photosynthetic efficiency (Fv/Fm)")+
  facet_grid(.~Day) +
  theme_classic() +
  theme(strip.text.x = element_text(size = 15),
        axis.text=element_text(size=15),
        axis.title=element_text(size=20), 
        legend.text=element_text(size=15), 
        legend.title=element_text(size=20))
        
ggsave(path = "plots",filename = "pam_heatwave_boxplot.png", width = 15, height = 7)
```

## Analysis of photosynthetic efficiency data

### Model 1: includes all morning measurement times

Testing assumptions of ANOVA:
```{r, message = F, warning = F}
#Filtering data to only include morning measurements
pam_morning_timeseries <- pam_clean %>% 
  filter(Date == "2021-11-01 06:00:00"| 
         Date == "2021-11-06 06:00:00"| 
         Date == "2021-11-08 06:00:00"|
         Date == "2021-11-09 06:00:00"| 
         Date =="2021-11-13 06:00:00") %>%
  mutate(Date = fct_relevel(as.factor(Date),
                            "2021-11-01 06:00:00",
                            "2021-11-06 06:00:00",
                            "2021-11-08 06:00:00",
                            "2021-11-09 06:00:00",
                            "2021-11-13 06:00:00"))

#Testing for normality and equal variances
shapiro_test(pam_morning_timeseries$Fv_Fm_av)
bartlett.test(Fv_Fm_av ~ Treatment, 
              data = pam_morning_timeseries)
#Results: Data is non-normal and does not have equal variances

#Testing transformations (log, arcsin, and square root)
pam_morning_timeseries <- pam_morning_timeseries %>%
  mutate(log_Fv_Fm = log(Fv_Fm_av), 
         sqrt_Fv_Fm = sqrt(Fv_Fm_av), 
         arc_Fv_Fm = asin(sqrt(Fv_Fm_av)))

#Testing assumptions of transformed data
shapiro_test(pam_morning_timeseries$log_Fv_Fm)
shapiro_test(pam_morning_timeseries$sqrt_Fv_Fm)
shapiro_test(pam_morning_timeseries$arc_Fv_Fm)
#Results: None of the transformations are normal. We will use a gamlss model rather than an ANOVA
```

Since the data does not fit the assumptions of an ANOVA, we will use a gamlss model:
```{r, echo=FALSE, message = F, warning = F}
#Finding distribution that best fits data
fitDist(Fv_Fm_av, data = pam_morning_timeseries, 
        type = "realAll", try.gamlss = T)
#Results: Gumbel is the best fit

#Visualizing distribution
histDist(pam_morning_timeseries$Fv_Fm_av, 
         "GU", 
         density = F, 
         main = "Gumbel")

#Full model
pam_morning_mod_full <- gamlss(Fv_Fm_av ~ Treatment + Date + Treatment*Date + random(Anemone_ID) + random(Site) + random(Bin), 
                               data = pam_morning_timeseries, 
                               family = GU(), 
                               control = gamlss.control(n.cyc=200))


#Backwards model selection to find best model:
pam_morning_mod_step <- stepGAIC(pam_morning_mod_full, 
                                 direction = "backward", 
                                 trace = F)
formula(pam_morning_mod_step)
summary(pam_morning_mod_step)
#Final model includes Treatment, Date, Treatment * Date, and random(Anemone_ID)

#Testing fit of full and reduced models using AIC
AIC(pam_morning_mod_full, 
    pam_morning_mod_step)
```

### Model 2: includes morning/afternoon measurements on first and third days of heatwave

Testing assumptions of ANOVA:
```{r message=FALSE, warning=FALSE}
#Filtering data to only include morning measurements
pam_heatwave_timeseries <- pam_clean %>% 
  filter(Date == "2021-11-06 06:00:00"| 
         Date == "2021-11-06 16:00:00"| 
         Date == "2021-11-08 06:00:00"|
         Date =="2021-11-08 16:00:00") %>%
  mutate(Date = fct_relevel(as.factor(Date),
                            "2021-11-06 06:00:00",
                            "2021-11-06 16:00:00",
                            "2021-11-08 06:00:00",
                            "2021-11-08 16:00:00"))

#Testing for normality and equal variances
shapiro_test(pam_heatwave_timeseries$Fv_Fm_av)
bartlett.test(Fv_Fm_av ~ Treatment,
              data = pam_heatwave_timeseries)
#Results: Data is non-normal and does not have equal variances

#Testing transformations (log, arcsin, and square root)
pam_heatwave_timeseries <- pam_heatwave_timeseries %>%
  mutate(log_Fv_Fm = log(Fv_Fm_av), 
         sqrt_Fv_Fm = sqrt(Fv_Fm_av), 
         arc_Fv_Fm = asin(sqrt(Fv_Fm_av)))

#Testing assumptions of transformed data
shapiro_test(pam_heatwave_timeseries$log_Fv_Fm)
shapiro_test(pam_heatwave_timeseries$sqrt_Fv_Fm)
shapiro_test(pam_heatwave_timeseries$arc_Fv_Fm)
#Results: None of the transformations are normal. We will use a gamlss model rather than an ANOVA
```
Since the data does not fit the assumptions of an ANOVA, we will use a gamlss model:
```{r, echo=FALSE, message = F, warning = F}
#Finding distribution that fits data
fitDist(Fv_Fm_av, data = pam_heatwave_timeseries, 
        type = "realAll", 
        try.gamlss = T)
#Best fit: generalized Gamma Loptatsidis-Green

#Visualizing distribution:
histDist(pam_heatwave_timeseries$Fv_Fm_av, 
         "GG", 
         density = F, 
         main = "generalised Gamma Lopatatsidis-Green")

#Full gamlss model
pam_heatwave_mod_full <- gamlss(Fv_Fm_av ~ Treatment + Date + Treatment*Date + random(as.factor(Bin)) + random(as.factor(Site)), 
                                data = pam_heatwave_timeseries, 
                                family = GG(), 
                                control = gamlss.control(n.cyc=200))

#Backwards model selection to find best model:
pam_heatwave_mod_final <- stepGAIC(pam_heatwave_mod_full, direction = "backward")
formula(pam_heatwave_mod_final)
summary(pam_heatwave_mod_final)
#Final model includes Treatment, Date, Treatment * Date, random(Bin), and random(Site)

#Testing fit of full and reduced models using AIC
AIC(pam_heatwave_mod_full, 
    pam_heatwave_mod_final)
```

# Base measurement analysis

## Plots

Boxplot of base measurements for each treatment and measurement time:
```{r, message = F, warning = F}
ggplot(base_clean, 
       aes(fill = Treatment, 
           x=Event,
           y= Average_Diameter)) + 
        theme_classic() +
         geom_boxplot() +
labs(x = "Measurement time", 
     y = "Base diameter (mm)") +
  scale_fill_manual(values =c('#89226AFF','#56B4E9FF','#E65154FF')) +
  theme(axis.text=element_text(size=15),
        axis.title=element_text(size=20), 
        legend.text=element_text(size=15), 
        legend.title=element_text(size=20))
ggsave(path = "plots",filename = "base_boxplot.png", width = 10, height = 7)
```

## Analyzing base diameter data

Testing assumptions for ANOVA:
```{r, message = F, warning = F}
shapiro_test(base_clean$Average_Diameter)
bartlett.test(Average_Diameter ~ Treatment, 
              data = base_clean)
#Data is non-normal and does not have equal variances

#Trying log transformation:
base_clean <- base_clean %>%
  mutate(log_diameter = log(Average_Diameter))

#Testing assumptions for log transformed data:
shapiro_test(base_clean$log_diameter)
bartlett.test(log_diameter ~ Treatment, 
              data = base_clean)
base_clean %>%
  group_by(Date, Treatment) %>%
  identify_outliers(log_diameter)
#Results: Log transformed data is normal and has equal variances. The data has one extreme 
#outlier, but this will not have a major effect on the results. We will use an two-way ANOVA 
#on the log transformed data.
```

Performing two-way ANOVA test:
```{r, message = F, warning = F}
#Two-way ANOVA on base diameter data with treatment and date as fixed effects, and 
#anemone ID as a random effect:
base_aov <- aov(log_diameter ~ Treatment*Date + random(Anemone_ID), 
                data=base_clean)
summary(base_aov)
TukeyHSD(base_aov)

#Testing fit of full and reduced models using AIC
AIC(base_aov)
```

# Feeding time

## Plots

Boxplot of feeding time data for each treatment and measurement time:
```{r, message = F, warning = F}
ggplot(food_clean, 
       aes(fill = Treatment, 
           x=Event, 
           y = Feeding_Time_Min)) + 
  theme_classic() +
  geom_boxplot() +
  labs(x = "Measurement time", 
       y = "Feeding time (min)") +
  scale_fill_manual(values =c('#89226AFF','#56B4E9FF','#E65154FF')) +
  theme(axis.text=element_text(size=15),
        axis.title=element_text(size=20), 
        legend.text=element_text(size=15), 
        legend.title=element_text(size=20))
ggsave(path = "plots",filename = "food_boxplot.png", width = 10, height = 7)
```

## Analyzing feeding time data

```{r, message = F, warning = F}
shapiro.test(food_clean$Feeding_Time_Min)
bartlett.test(Feeding_Time_Min ~ Treatment, 
              data = food_clean)
#Data is non-normal and does not have equal variances

#Trying log transformation:
food_clean<-food_clean%>%
  mutate(log_Feeding_Time_Min = log(Feeding_Time_Min))

shapiro_test(food_clean$log_Feeding_Time_Min)
bartlett.test(log_Feeding_Time_Min ~ Treatment, 
              data = food_clean)
food_clean %>%
  group_by(Date, Treatment) %>%
  identify_outliers(log_Feeding_Time_Min)
#Log transformed data fits normal distribution and has equal variances. There are also no extreme outliers. 
#We will use a two-way ANOVA to analyze the log-transformed data.
```

Performing two-way ANOVA on log transformed feeding time data:
```{r, message = F, warning = F}
#Two-way anova with treatment and date as fixed effects, and anemone ID as a random effect
food_aov <- aov(log_Feeding_Time_Min ~ Treatment*Date + random(Anemone_ID), 
                data=food_clean)
summary(food_aov)
TukeyHSD(food_aov)

#Testing fit of full and reduced models using AIC
AIC(food_aov)
```

# Heatwave response data analysis

## Plots

Creating a stacked bar plot showing proportions of open, closed, and partially open anemones at each hour of the heatwave on each day
```{r, message = F, warning = F}
ggplot(data = open_closed_summary, 
       aes(x = Time_Block, 
           y = n, 
           fill = Open_Closed)) + 
  geom_bar(position = "fill", 
           stat = "identity") + 
  facet_grid(Event ~ Treatment) +
  labs(x="Time (hours)", 
       y = "Proportion", 
       fill = "Behaviour") +
  theme_classic()+
scale_fill_manual(values = c('#FFE082','#2C728EFF','#E65154FF'))+
  theme(strip.text.x = element_text(size = 15),
        strip.text.y = element_text(size= 15),
        axis.text=element_text(size=15),
        axis.title=element_text(size=20), 
        legend.text=element_text(size=15), 
        legend.title=element_text(size=20))
ggsave(path = "plots",filename = "open_closed_plot.png", width = 10, height = 7)
```

## Analyzing data
Exploratory data analysis:
```{r, message = F, warning = F}
#Summarizing the data
summary(open_closed_summary)
#Making frequency table
table(open_closed_clean$Treatment, 
      open_closed_clean$Open_Closed)

```


Ordinal regression model (unsuccessful)
```{r, echo = F message = F, warning = F}
ord_model = clmm(Open_Closed ~ Treatment:Time_Block:Date + (1|Anemone_ID), 
                 data = open_closed_clean)
summary(ord_model)
#This technique did not accurately model our data due to the low frequencies of measurements 
#in some categories
```

Bayesian regression analysis
```{r, message = F, warning = F}
#Running a Bayesian model with weakly flat priors. 
#Fixed effects are treatment, time block, and date. 
#Anemone ID is a random effect.

#Run this code once before running model
#options(mc.cores=parallel::detectCores())

#bay_mod <- brm(Open_Closed ~ Treatment + Time_Block + Date + (1|Anemone_ID), 
               #data = open_closed_clean, 
               #family = cumulative("logit"))
#summary(bay_mod)

#Calculating percent confidence for each treatment, date, and time block
#response_post = posterior_samples(bay_mod)
#sum(response_post$b_Treatment30C > 0) / 4000 # 0.82275
#sum(response_post$b_Treatment25C > 0) / 4000 # 0.2765
#sum(response_post$b_Time_Block1 > 0) / 4000 # 0.888
#sum(response_post$b_Time_Block2 > 0) / 4000 #0.9975
#sum(response_post$b_Time_Block3 > 0) / 4000 # 0.88
#sum(response_post$b_Time_Block4 > 0) / 4000 # 0.99875
#sum(response_post$b_Time_Block5 > 0) / 4000 # 1
#sum(response_post$b_Time_Block6 > 0) / 4000 # 1
#sum(response_post$b_Date11D7D2021 > 0) / 4000 # 0.17525
#sum(response_post$b_Date11D8D2021 > 0) / 4000 # 0.42775
```

# Hemocytometer data analysis

## Plots

Boxplots of dinoflagellate density and mitotic index at each measurement time:
```{r, message = F, warning = F}
#Dinoflagellate density
p1 <- ggplot(data = hemo_summary, 
       aes(x= Date, 
           y = mean_Dino_Density, 
           group = Treatment, 
           colour = Treatment)) +
  theme_classic() +
  geom_errorbar(aes(ymin = mean_Dino_Density - se_Dino_Density, 
                    ymax = mean_Dino_Density + se_Dino_Density), 
                    width=30000) +
  geom_vline(xintercept = as.POSIXct("2021-11-06 09:00:00"),
             linetype = "dotted", 
             size = 1) +
  geom_vline(xintercept = as.POSIXct("2021-11-08 16:00:00"), 
             linetype = "dotted", 
             size = 1)+
  geom_point(size=3) +
  geom_line(lwd = 2) +
  scale_fill_manual(values =c('#89226AFF','#56B4E9FF','#E65154FF')) +
  scale_colour_manual(values = c('#89226AFF','#56B4E9FF','#E65154FF'))+
  labs(x="Date", 
       y = "Zooxanthellae density (cells/mg)")+
  scale_x_datetime(breaks=date_breaks("3 days"),
                    labels=date_format("%b-%d")) +
  theme(legend.position = "none")

#Mitotic Index of Dinoflagellates
p2 = ggplot(data = hemo_summary, 
       aes(x= Date, 
           y = mean_Dino_MI, 
           group = Treatment, 
           colour = Treatment)) +
  theme_classic() +
  geom_errorbar(aes(ymin = mean_Dino_MI - se_Dino_MI, 
                    ymax = mean_Dino_MI + se_Dino_MI), 
                    width=30000) +
  geom_point(size=3) +
  geom_line(lwd = 2) +
  geom_vline(xintercept = as.POSIXct("2021-11-06 09:00:00"),
             linetype = "dotted", 
             size = 1) +
  geom_vline(xintercept = as.POSIXct("2021-11-08 16:00:00"), 
             linetype = "dotted", 
             size = 1)+
   scale_fill_manual(values = c('#89226AFF','#56B4E9FF','#E65154FF')) +
  scale_colour_manual(values = c('#89226AFF','#56B4E9FF','#E65154FF'))+
  labs(x="Date", 
       y = "Zooxanthellae Mitotic Index")+
   scale_x_datetime(breaks=date_breaks("3 days"),
                    labels=date_format("%b-%d")) 

#Combining plots
p1 + p2

ggsave(path = "plots",filename = "dinoflagellate_density_MI.png", width = 10, height = 4)
```

## Analyzing zooxanthellae density and mitotic index

### Zooxanthellae density
```{r, message = F, warning = F}
#Dinoflagellate density
shapiro_test(hemo_clean$Dino_Density)
bartlett.test(Dino_Density ~ Treatment, 
              data = hemo_clean)
#Data has equal variances but is not normal

#log transformation:
hemo_clean <- hemo_clean %>%
  mutate(log_Dino_Density = log(Dino_Density))

shapiro_test(hemo_clean$log_Dino_Density)
bartlett.test(log_Dino_Density ~ Treatment, 
              data = hemo_clean)
hemo_clean %>%
  group_by(Treatment, Date) %>%
  identify_outliers(log_Dino_Density) 
#Log transformed data is normal and has equal variances. 
#There are two extreme outliers but this will not have a major impact on the results. 
#We will use a two-way ANOVA to analyze this data.
```

Two-way AVOVA on zooxanthellae density data:
```{r, message = F, warning = F}
Dino_Density_aov <- aov(log_Dino_Density ~ Treatment*as.factor(Date) + random(Anemone_ID),
                        data=hemo_clean)
summary(Dino_Density_aov)
TukeyHSD(Dino_Density_aov)

#Testing fit of full and reduced models using AIC
AIC(Dino_Density_aov)
```


### Zooxanthellae mitotic index
```{r}
#Dinoflagellate mitotic index
shapiro_test(hemo_clean$Dino_MI)
bartlett.test(Dino_MI ~ Treatment, 
              data = hemo_clean)
#Data has equal variances but is not normal

# Trying transformations:
hemo_clean <- hemo_clean %>%
  mutate(log_Dino_MI = log(Dino_MI + 0.1), 
         arcsine_Dino_MI = asin(sqrt(Dino_MI)), 
         sqrt_Dino_MI = sqrt(Dino_MI))

# Testing normality of transformed data
shapiro_test(hemo_clean$log_Dino_MI) 
shapiro_test(hemo_clean$arcsine_Dino_MI) 
shapiro_test(hemo_clean$sqrt_Dino_MI) 
#Arcsin and square root tranformations are not normal, but the log transformed data is.

bartlett.test(log_Dino_MI ~ Treatment, 
              data = hemo_clean) #0.03211
#log transformed data has equal variances

hemo_clean %>%
  group_by(Treatment, Date) %>%
  identify_outliers(log_Dino_MI) 
#The data has three extreme outliers, but this will not have a major effect on the results. 
#We will use an two-way ANOVA on the log transformed data.
```

Two-way ANOVA on mitotic index data:
```{r, message = F, warning = F}
#Dinoflagellate mitotic index
Dino_MI_aov <- aov(log_Dino_MI ~ Treatment*as.factor(Date) + random(Anemone_ID), 
                   data=hemo_clean)
summary(Dino_MI_aov)
TukeyHSD(Dino_MI_aov)

#Testing fit of full and reduced models using AIC
AIC(Dino_MI_aov)
```

# Heatwave temperature

##Plots

Creating a plot to show the average temperature in the final 3 hours of the heatwave (after the temperature ramp) on each day
```{r, message = F, warning = F}
ggplot(data = temp_summary,
       aes(x=Time_Block, 
           y = mean_temp,
           group=Treatment,
           colour = Treatment))+
  theme_classic() +
  geom_point(size=2.5) +
  geom_line(lwd = 1.5) +
   facet_grid(. ~ Event) +
  scale_fill_manual(values = c('#89226AFF','#56B4E9FF','#E65154FF')) +
  scale_colour_manual(values = c('#89226AFF','#56B4E9FF','#E65154FF'))+
  labs(x="Time (Hours)", 
       y ="Temperature (°C)") +
  theme(strip.text.y = element_text(size = 15),
       axis.text=element_text(size=15),
       axis.title=element_text(size=20), 
       legend.text=element_text(size=15), 
       legend.title=element_text(size=20))
  
ggsave(path = "plots",filename = "temp_plot.png", width = 10, height = 7)
```